---
title: "Clustering-gam-data"
author: "Ruiman"
date: "2024-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All simulated datasets are included in the folder \textcolor{red}{Sim-Data}. Here, we take the shifted Doppler signals with b-splines basis functions as an example to illustrate the usage of the Fclust-RST method. 

- We first install and load the required R packages:
```{r echo=TRUE, message=FALSE, warning=FALSE}
rm(list = ls())
packages = c( "gridExtra", "deldir", "fields", "igraph",  "devtools", "ggplot2", "Matrix.utils", "grid", 
             "parallel", "class", "spgwr", 'sf',
             "salso", 'bsfc','cccd',
             "fda","mclust")

## Now load or install&load all packages
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      # install.packages(x, dependencies = TRUE, 
      #          INSTALL_opts = '--no-lock', 
      #          repos = "http://cran.us.r-project.org")
      library(x, character.only = TRUE)
    }else
      library(x, character.only = TRUE)
  }
)
path_proj <- '~/Documents/project3'
path_data <- file.path(path_proj, "data")
path_raw <- file.path(path_data, "raw")
path_sim <- file.path(path_data, "simulated")
path_im <- file.path(path_data,  "image")
```
Load spatial-temporal data and map

```{r echo=TRUE, message=FALSE, warning=FALSE}
##Load data
data = readRDS(file.path(path_sim, 'spatial-data-gam.rds'))
map <-  readRDS(file.path(path_raw, "spatial-domain-clusters.rds"))
cluster_true = data$cluster
eta = data$eta
Y = data$Y
nt=dim(Y)[1]; ns=dim(Y)[2]  
## build design matrix
time = seq(0, 1, length.out = nt)
# compute fixed effects
X = poly(time, 2, raw = TRUE, simple = TRUE)
population <- matrix(map$population, nrow = nt, ncol = ns, byrow = T)
```

- Obtain initial estimates for the between-cluster parameters. 
   \begin{itemize}
     \item  The hyperparameter $\mathcal{G}$ is fixed to be the true graph $\mathcal{G}_0$ used for generating the cluster-wise constant Doppler signals.      \item  The initial values of wavelet coefficients, $\{\hat{\boldsymbol{\beta}}(\mathbf{s})\}$, are obtained by using the least-square estimates for each location $\mathbf{s}$. 
     \item  For each edge of $\mathcal{G}$ connecting $\mathbf{s}$ and $\mathbf{s}^{\prime}$, the Euclidean distance of $\hat{\boldsymbol{\beta}}(\mathbf{s})$ and $\hat{\boldsymbol{\beta}}(\mathbf{s}^{\prime})$ is used as its edge weight. Then, the initial minimum spanning tree (MST), $\mathcal{T}^{(0)}$, is obtained based on $\mathcal{G}$ and its assigned edge weights. 
     \item  By removing $(k_p^{(0)}-1)$ edges of $\mathcal{T}^{(0)}$, we obtain $k_p^{(0)}$ initial clusters. 
  \end{itemize}

We could use random weights to construct spanning tree. 
```{r warning=FALSE}
geodata = initial.mst.build(map, method = 'knn', para = 10, nclust = 10)
plot(geodata$mstgraph.ini)
```
To guarantee spatial continuity, we want to use adjacent matrix to build the graph and MST. A better initial weight is also provided.
```{r echo=TRUE, warning=FALSE}
seed = 81
graph0 <- ConstructGraph0(map,method='adjmat', seed = seed)
ratio = matrix(0,nrow = nt, ncol = ns)
for(i in 1:ns){ratio[,i] = as.matrix(Y[,i]/population[,i])}
# prepare initial weights based on observations
beta.ini= ginv(t(X) %*% X) %*% t(X) %*% ratio # p * s
beta.ini = t(beta.ini) 
inc.mat=get.edgelist(graph0, names = F)
weights=sqrt(rowSums((beta.ini[inc.mat[,1],]-beta.ini[inc.mat[,2],])^2))
# Build MST 

geodata = initial.mst.build(map,method='adjmat',nclust = 10, weights = weights, seed = seed)
geodata$cluster
cluster_true
adjustedRandIndex(geodata$cluster,cluster_true)
table(geodata$cluster,cluster_true)
```
- Initialize other model parameters:
 hyperpar is a parameter contains the penalty of the number of clusters
 formula is for fitting within-cluster model, thus, the response name should be 'Yk' and the covariate should be named as 'Xk'
- Run the Fclust-RST method on normal response data:
```{r eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
path_res <- file.path(path_proj, "data", "result","normal_gam_exp.rds")
formula= Yk~ f(Xk, model = "rw1", scale.model = T)
system.time({
bsfc(eta, graphdata = geodata, X = time, N = NULL,
                formula, family = "normal", hyperpar = list(c = 0.5),
                correction = T, niter = 500, burnin = 0, thin = 1, path_save = path_res) 
 })
```
If more iterations need to converge, we could use continue_bsfc() 
```{r eval=FALSE, include=FALSE}
result = readRDS(path_res)
continue_bsfc(result, eta, X = time, N = NULL, graph = graph0,
                formula, family = "normal", hyperpar = list(c = 0.1),
                correction = T, niter = 5000, burnin = 1000, thin = 1, path_save = path_res) 
```
```{r}
result = readRDS(path_res)
clust_res = result$cluster[nrow(result$cluster),]
cluster_true
table(cluster_true,clust_res)
adjustedRandIndex(clust_res, cluster_true)
p = max(clust_res)
final_model =  result$model

plotClusterMap(cluster_true = cluster_true,clust_res = as.factor(clust_res), map, file.path(path_im, 'Cluster_map_gam_normal_res.png'))

plotClusterFun(eta,nt,clust_res,final_model,a = 3, b=4, file.path(path_im, 'Cluster_fun_gam_normal_res.png'))


  data.frame(id = 1:length(result$log_mlike), log_mlike = result$log_mlike) |>
    ggplot() +
    geom_line(aes(id, log_mlike)) +
    labs(x = "iteration", y = "log marginal likelihood")
table(cluster_true, result$cluster[nrow(result$cluster),])
```
If more iterations need to converge, we could use continue_bsfc() 
```{r eval=FALSE, include=FALSE}
result = readRDS(path_res)
formula= Yk~ f(Xk, model = "rw1", scale.model = T) + f(id, model = "iid")
path_res_con <- file.path(path_proj, "data", "result","bsfc_gam_poisson_con.rds")
continue_bsfc(result, Y, X = time, N = population, graph = graph0,
                formula, family = "poisson", hyperpar = list(c = 0.1),
                correction = T, niter = 1000, burnin =0, thin = 1, path_save = path_res) 
```

```{r}
path_res <- file.path(path_proj, "data", "result","bsfc_gam_poisson.rds")
result = readRDS(path_res)

clust_res = result$cluster[nrow(result$cluster),]
cluster_true
table(cluster_true,clust_res)
adjustedRandIndex(clust_res, cluster_true)
p = max(clust_res)
final_model =  result$model

plotClusterMap(cluster_true = cluster_true,clust_res = as.factor(clust_res), map, file.path(path_im, 'Cluster_map_gam_poisson_res.png'))

plotClusterFun(exp(eta),nt,clust_res,final_model,a = 3, b=4, file.path(path_im, 'Cluster_fun_gam_poisson_res.png'))


  data.frame(id = 1:length(result$log_mlike), log_mlike = result$log_mlike) |>
    ggplot() +
    geom_line(aes(id, log_mlike)) +
    labs(x = "iteration", y = "log marginal likelihood")
table(cluster_true, result$cluster[nrow(result$cluster),])
```


