
---
title: "Creating Custom Initial Graphs"
author: "Ruiman Zhong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom Initial Graphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The `sfclust` package currently supports the use of adjacency matrices for graph creation. However, users who need more advanced or customized graphs can still create them using external libraries and pass these graphs into the clustering functions. This vignette demonstrates how to use other methods like k-nearest neighbors, relative neighborhood graphs, and Delaunay triangulation to create custom graphs.

# Available Graph Construction Methods

### 1. K-Nearest Neighbors (KNN)
Users can create a k-nearest neighbors (KNN) graph using the `nng()` function from the `cccd` package or write a custom function. Below is an example using the `cccd` package:

```{r}
library(cccd)
library(sfclust)
library(sf)
library(ggraph)
data("us_covid")
map <-  us_covid$map
coords <- st_coordinates(st_centroid(map))  # Get coordinates from spatial object

# Create KNN graph
graph_knn <- nng(coords, k = 3)
plotGraph(map,coords,graph_knn, title = NULL)
```

### 2. KNN with Geographical Distance
In cases where geographical distances are important, the following code shows how to use `rdist.earth()` from the `fields` package:

```{r}
library(fields)
coords <- st_coordinates(st_centroid(map))

# KNN with geographical distance
dist_geo <- rdist.earth(coords, miles = FALSE)
graph_knn_geo <- nng(coords, k = 3)
plotGraph(map,coords,graph_knn_geo, title = NULL)
```

### 3. Custom Graph from Adjacency Matrix
The package natively supports creating graphs from adjacency matrices. Here is a reminder of how to create a simple adjacency matrix and graph:

```{r}
library(spdep)
neighbors <- poly2nb(map)
adjmat <- nb2mat(neighbors, style = "B", zero.policy = TRUE)
graph_adjmat <- graph_from_adjacency_matrix(adjmat, mode = "undirected")
plotGraph(map,coords,graph_adjmat, title = NULL)
```
# Create MST and initial clusters
```{r}
adj_matrix_knn <- as_adjacency_matrix(graph_knn, sparse = FALSE)

geodata <- initial_cluster(adj_matrix_knn, nclust = 5)
```
# Passing Custom Graphs to the `sfclust` Package

Once you have created the initial parameters based on your custom graph, you can pass it into the package’s functions like so:

```{r eval=FALSE}
formula <- Y ~ 1 + f(idt, model = "ar1")
# Assuming 'graph_custom' is the graph you've created:
result <- sfclust(Y = us_covid$Y, graphdata = geodata, formula = formula, niter = 1000)
```

This approach allows you to control how the initial graph is constructed, enabling flexibility for specific use cases or datasets.

# Conclusion

While the default graph creation method in the package is adjacency matrices, users can still leverage external tools and libraries to create custom graphs, giving them greater control over how their spatial data is structured. 
