---
title: "Spatial_functional_clustering_for_glm_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial_functional_clustering_for_Poisson_glm_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All simulated datasets are included in the folder \textcolor{red}{Sim-Data}. Here, we take the shifted Doppler signals with b-splines basis functions as an example to illustrate the usage of the Fclust-RST method. 

- We first install and load the required R packages:
```{r echo=TRUE, message=FALSE, warning=FALSE}
rm(list = ls())
packages = c( "gridExtra", "deldir", "fields", "igraph",  "devtools", "ggplot2", "Matrix.utils", "grid", 
             "parallel", "class", "spgwr", 'sf',
             "salso", 'bsfc',
             "fda","mclust")

## Now load or install&load all packages
package.check <- lapply(
  packages,
  FUN = function(x) {library(x, character.only = TRUE)})
path_proj <- '~/Documents/project3'
path_data <- file.path(path_proj, "data")
path_raw <- file.path(path_data, "raw")
path_sim <- file.path(path_data, "simulated")
path_res <- file.path(path_proj, "data", "result","bsfc-glm-poisson.rds")
path_im <- file.path(path_data,  "image")
##Load data
data = readRDS(file.path(path_sim, 'spatial-data-glm.rds'))
map <-  readRDS(file.path(path_raw, "spatial-domain-clusters.rds"))
cluster_true = data$cluster
eta = data$eta
Y = data$Y
nt=dim(Y)[1]; ns=dim(Y)[2]  
## build design matrix
time = seq(0, 1, length.out = nt)
# compute fixed effects
X = poly(time, 2, raw = TRUE, simple = TRUE)
population <- matrix(map$population, nrow = nt, ncol = ns, byrow = T)
```

- Obtain initial estimates for the between-cluster parameters. 
   \begin{itemize}
     \item  The hyperparameter $\mathcal{G}$ is fixed to be the true graph $\mathcal{G}_0$ used for generating the cluster-wise constant Doppler signals.      \item  The initial values of wavelet coefficients, $\{\hat{\boldsymbol{\beta}}(\mathbf{s})\}$, are obtained by using the least-square estimates for each location $\mathbf{s}$. 
     \item  For each edge of $\mathcal{G}$ connecting $\mathbf{s}$ and $\mathbf{s}^{\prime}$, the Euclidean distance of $\hat{\boldsymbol{\beta}}(\mathbf{s})$ and $\hat{\boldsymbol{\beta}}(\mathbf{s}^{\prime})$ is used as its edge weight. Then, the initial minimum spanning tree (MST), $\mathcal{T}^{(0)}$, is obtained based on $\mathcal{G}$ and its assigned edge weights. 
     \item  By removing $(k_p^{(0)}-1)$ edges of $\mathcal{T}^{(0)}$, we obtain $k_p^{(0)}$ initial clusters. 
  \end{itemize}
```{r echo=TRUE, warning=FALSE}
# Build MST 
k =15
seed = 43
coords <- st_coordinates(st_centroid(map)[,3])
# create graph
geodata = initial.mst.build(map,method='adjmat',nclust = k, weights = NULL, seed = seed)

formula= Yk~ 1+ Xk + f(id, model = 'iid')

bsfc(Y, graphdata = geodata, X = X, N = population,
                formula, family = "poisson", hyperpar = list(c =  0.9498706),
                correction = F, niter = 2000, burnin = 0, thin = 1, path_save = path_res) 
```

- Plot the clustering results:

```{r echo=TRUE, message=FALSE}

#####Load the results of Fclust-RST
result = readRDS(path_res)
cluster <- result$cluster[1500:2000,]
pihat <- salso::psm(cluster)
clust_idx=which.min(salso::binder(cluster, pihat))
clust_res=cluster[clust_idx,]
###Obtain the adjusted Rand index with the true partition
adjustedRandIndex(clust_res, cluster_true)

p = max(clust_res)
final_model =  lapply(1:p,log_mlik_each, Y, clust_res, X, N = NULL, formula, family = "poisson", correction = FALSE, detailed = T)

plotClusterMap(cluster_true = cluster_true,clust_res = as.factor(clust_res), map,height = 4, file.path(path_im, 'map_glm_poisson_res.jpg'))
plotClusterIter(result$cluster, height = 8, filepath = file.path(path_im, 'iter_glm.jpg'))
plotmlikeIter(result$log_mlike, height = 8, filepath = file.path(path_im, 'mlik_glm.jpg'))

```

### test
beta = cbind(
    c(1, 0), c(-1, 0), c(0, 0), c(-3, 3), c(3, -3),
    c(1, 0), c(-1, 0), c(0, 0), c(-3, 3), c(3, -3)
)

